---
title: "Column types"
author: "Hadley Wickham"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Column types}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
library(readr)
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Column types

Currently, readr automatically recognises the following types of columns:

* `col_logical()` [l], containing only `T`, `F`, `TRUE` or `FALSE`.

* `col_integer()` [i], integers.

* `col_double()` [d], doubles.

* `col_character()` [c], everything else.

* `col_date(format = "")` [D]: with the locale's `date_format`.

* `col_time(format = "")` [t]: with the locale's `time_format`.

* `col_datetime(format = "")` [T]: ISO8601 date times

* `col_number()` [n], numbers containing the `grouping_mark`

To recognise these columns, readr inspects the first 1000 rows of your dataset. This is not guaranteed to be perfect, but it's fast and a reasonable heuristic. If you get a lot of parsing failures, you'll need to re-read the file, either increasing `guess_max` to overriding the default choices as described below.

You can also manually specify other column types:

* `col_skip()` [_, -], don't import this column.

* `col_date(format)`, dates with given format.

* `col_datetime(format, tz)`, date times with given format. If the timezone
  is UTC (the default), this is >20x faster than loading then parsing with
  `strptime()`.

* `col_time(format)`, times. Returned as number of seconds past midnight.

* `col_factor(levels, ordered)`, parse a fixed set of known values into a
  factor

Use the `col_types` argument to override the default choices. There are two ways to use it:

* With a string: `"dc__d"`: read first column as double, second as character,
  skip the next two and read the last column as a double. (There's no way to
  use this form with types that take additional parameters.)

* With a (named) list of col objects:

    ```R
    read_csv("iris.csv", col_types = cols(
      Sepal.Length = col_double(),
      Sepal.Width = col_double(),
      Petal.Length = col_double(),
      Petal.Width = col_double(),
      Species = col_factor(c("setosa", "versicolor", "virginica"))
    ))
    ```

    Or, with their abbreviations:

    ```R
      read_csv("iris.csv", col_types = cols(
      Sepal.Length = "d",
      Sepal.Width = "d",
      Petal.Length = "d",
      Petal.Width = "d",
      Species = col_factor(c("setosa", "versicolor", "virginica"))
    ))
    ```

    Any omitted columns will be parsed automatically, so the previous call
    will lead to the same result as:

    ```R
    read_csv("iris.csv", col_types = cols(
      Species = col_factor(c("setosa", "versicolor", "virginica")))
    )
    ```

    You can also set a default type that will be used instead of
    relying on the automatic detection for columns you don't specify:

    ```R
    read_csv("iris.csv", col_types = cols(
      Species = col_factor(c("setosa", "versicolor", "virginica")),
      .default = col_double())
    )
    ```

    If you only want to read specified columns, use `cols_only()`:

    ```R
    read_csv("iris.csv", col_types = cols_only(
      Species = col_factor(c("setosa", "versicolor", "virginica")))
    )
    ```

When reading files interactively the first 20 rows of the `col_spec()` used are
printed. `option(readr.num_columns)` can be used to change the number of
columns to be printed, setting the value to 0 disables printing.

readr attaches the spec used for the file to the output. It can be retrieved by
calling `spec()` on the object.

```{r}
data <- read_csv(readr_example("mtcars.csv"))
data

# Every table returned has a spec attribute
s <- spec(data)
s

# Alternatively you can use a spec function instead, which will only read the
# first 1000 rows (user configurable with guess_max)
s <- spec_csv(readr_example("mtcars.csv"))
s

# Automatically set the default to the most common type
cols_condense(s)

# If the spec has a default of skip then uses cols_only
s$default <- col_skip()
s

# Otherwise set the default to the proper type
s$default <- col_character()
s

# The print method takes a n parameter to return only that number of columns
print(s, n = 5)

# When reading this is set to 20 by default, set 
# options("readr.num_columns" = x) to change
options("readr.num_columns" = 5)
data <- read_csv(readr_example("mtcars.csv"))

# Setting it to 0 disables printing
options("readr.num_columns" = 0)
data <- read_csv(readr_example("mtcars.csv"))
```

### Spec types

## Column parsers

As well as specifying how to parse a column from a file on disk, each of the `col_xyz()` functions has an equivalent `parse_xyz()` that parsers a character vector. These are useful for testing and examples, and for rapidly experimenting to figure out how to parse a vector given a few examples.

### Base types

`parse_logical()`, `parse_integer()`, `parse_double()`, and `parse_character()` are straightforward parsers that produce the corresponding atomic vector.

Make sure to read `vignette("locales")` to learn how to deal with doubles.

### Numbers

`parse_integer()` and `parse_double()` are strict: the input string must be a single number with no leading or trailing characters. `parse_number()` is more flexible: it ignores non-numeric prefixes and suffixes, and knows how to deal with grouping marks. This makes it suitable for reading currencies and percentages:

```{r}
parse_number(c("0%", "10%", "150%"))
parse_number(c("$1,234.5", "$12.45"))
```

Note that `guess_parser()` will only guess that a string is a number if it has no leading or trailing characters (after trimming whitespace), otherwise it's too prone to false positives. That means you'll typically needed to explicitly supply the column type for number columns.

```{r}
guess_parser("$1,234")
guess_parser("1,234")
```

### Date times

readr supports three types of date/time data:

* dates: number of days since 1970-01-01.
* times: number of seconds since midnight.
* datetimes: number of seconds since midnight 1970-01-01.

readr will guess date time fields if they're in ISO8601 format:

```{r}
parse_datetime("2010-10-01 21:45")
parse_date("2010-10-01")
```

Otherwise, you'll need to specify the format yourself:

```{r}
parse_datetime("1 January, 2010", "%d %B, %Y")
parse_datetime("02/02/15", "%m/%d/%y")
```

### Factors

When reading a column that has a known set of values, you can read directly into a factor.

```{r}
parse_factor(c("a", "b", "a"), levels = c("a", "b", "c"))
```

readr will never turn a character vector into a factor unless you explicitly ask for it.
